---
title: "The effect of mask-wearing & hand-washing during COVID-19 pandemic"
author: "Sin-Yi Chou"
date: "2020-05-09"
output: html_document
---

```{r setup, include=F}
library(knitr)
library(kableExtra)
library(dplyr)
library(tibble)
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggsci)
library(stringr)
library(lubridate)
library(plotly)
library(readxl)
library(scales)

```

# COVID-19 in Taiwan 

```{r, echo=F,warning=F,fig.align='center', fig.width=9, fig.height= 7.5}

df_tw_covid_case = read.csv("data/taiwan_cdc/covid19/嚴重特殊傳染性肺炎TaiwanCDC公布病例資訊-確診病例.csv", stringsAsFactors = F)

clean_df_tw_covid_case = 
  df_tw_covid_case %>% 
  select(case_no = `案例編號`,
         onset_date = `確診日`, 
         domestic_index = `境外或是本土`) %>%
  mutate(domestic_index = case_when( domestic_index =="境外" ~ "Imported",
                                     domestic_index == "本土" ~ "Indigenous",
                                     TRUE ~ "Dunmu fleet (Navy)") 
         ) %>% 
  mutate(onset_date = ymd(onset_date)) %>% 
  group_by(onset_date, domestic_index) %>% 
  summarise(case = n()) %>% 
  ungroup() %>% 
  spread(key=domestic_index, value = case) %>% 
  replace(is.na(.),0) %>% 
  arrange(onset_date) %>% 
  mutate(cum_indigenous = cumsum(Indigenous),
         cum_imported = cumsum(Imported),
         cum_dunmu= cumsum(`Dunmu fleet (Navy)`)) %>%
  mutate(total = cum_indigenous + cum_imported + cum_dunmu) %>% 
  gather(key = "index", value="case", -onset_date) %>% 
  mutate(cum_index = case_when(str_starts(index,"cum") ~  "cumulative",
                               index == "total" ~ "total",
                               TRUE ~ "daily")) %>% 
  mutate(index = case_when(index =="cum_imported" ~ "Imported",
                           index == "cum_indigenous" ~ "Indigenous",
                           index == "cum_dunmu" ~ "Dunmu fleet (Navy)",
                           TRUE ~ index)) %>% 
  mutate(index = factor(index , levels = c("Indigenous","Imported","Dunmu fleet (Navy)" ,"total")))


a=clean_df_tw_covid_case %>% 
  filter(index!="total") %>%
  ggplot() +
  geom_bar(aes(x=onset_date, y=case, fill=index),
           stat="identity", position="stack",
           data = clean_df_tw_covid_case %>% 
             filter(cum_index =="daily" )
             )+
  geom_point(aes(x=onset_date, y=case, color=index),
           data = clean_df_tw_covid_case %>% 
             filter(cum_index =="cumulative"))+
  geom_line(aes(x=onset_date, y=case, color=index),
             data = clean_df_tw_covid_case %>% 
               filter(cum_index =="cumulative"))+
  labs(x=NULL, y='Confirmed cases', 
       title = "Daily & cumulative confirmed COVID-19 cases by sources in Taiwan",
       fill=NULL, color=NULL)+
  facet_wrap(vars(index), ncol = 1)+
  scale_x_date(labels = label_date_short(), breaks = pretty_breaks(10))+
  scale_fill_jco()+
  scale_color_jco()+
  theme_bw()+
    theme(text=element_text(size=18),
        plot.title = element_text(size=15),
        plot.subtitle = element_text(size=10)
  )


plot = ggplotly(a) 
plot[['x']][['layout']][['annotations']][[1]] <- c()
plot[['x']][['data']][[1]]$name = "Indigenous"
plot[['x']][['data']][[1]]$legendgroup = "Indigenous"
plot[['x']][['data']][[2]]$name = "Imported"
plot[['x']][['data']][[2]]$legendgroup = "Imported"
plot[['x']][['data']][[3]]$name = "Dunmu fleet (Navy)"
plot[['x']][['data']][[3]]$legendgroup = "Dunmu fleet (Navy)"
plot[['x']][['data']][[4]]$name = "Indigenous"
plot[['x']][['data']][[4]]$legendgroup = "Indigenous"
plot[['x']][['data']][[5]]$name = "Imported"
plot[['x']][['data']][[5]]$legendgroup = "Imported"
plot[['x']][['data']][[6]]$name = "Dunmu fleet (Navy)"
plot[['x']][['data']][[6]]$legendgroup = "Dunmu fleet (Navy)"

plot %>% 
 layout(font=list(family="helvetica"),
         legend = list(x = 100, y = 0.6, 
                       title=list(text='Source'),
                       orientation = 'v'),
         title = list(
           font = list(
             size = 20
           ), 
           y=3,
           text = "Cumulative & daily confirmed COVID-19 cases by sources in Taiwan"
         ),
        annotations = list(
          list(text="Confirmed cases",
                           x=-0.07,
                           y=0.5,
                           ax=-0,
                           ay=0,
                           showarrow=F,
                           font = list(size=18),
                           xref="paper",
                           yref="paper",
                           textangle=-90,
                           xanchor="right",
                           yanchor="center",
                           annotationType="axis"
                           ),
                     list(x = 1.05, y = -0.11, text = "Data source: Taiwan CDC", 
                showarrow = F, xref='paper', yref='paper', 
                xanchor='right', yanchor='auto', xshift=-15, yshift=-30,
                font=list(size=13, color="grey 10"))),
        margin = list(
          t=80, b=0 
        )) 

```

Note: Dummu fleet is a cluster of cases from a navy ship that returned to Taiwan after missions overseas 

# Other infectious diseases in Taiwan 

## Influenza 

### Influenza with severe complications 
<!-- ## 全國流感併發重症本土病例及境外移入病例 -->

```{r, echo=F, fig.align='center', fig.width=9}

df_severe_flu = read.csv("data/taiwan_cdc/全國流感併發重症本土病例及境外移入病例同期比較趨勢圖_all2.csv", stringsAsFactors = F)


list_new_col = df_severe_flu %>% colnames()

list_new_col[1]="onset_month"

list_new_col = str_replace_all(list_new_col, "X","")

colnames(df_severe_flu) = list_new_col

clean_df_severe_flu = 
  df_severe_flu %>% 
  gather(key="year", value="case", -onset_month) %>% 
  mutate(flu_season = if_else(onset_month>=6 & onset_month<=9,
                              "not flu season",
                              "flu season")) %>% 
  mutate(month= factor(month.abb[onset_month], levels = month.abb, ordered = T)) %>%
  mutate(plot_date = ymd(str_c(year,"-",month,"-","15"))) %>% 
  mutate(this_year = if_else(year==2020, "this year", "other years")) %>% 
  mutate(year2=year) %>% 
  filter(!(year==2020 & onset_month >=5))

a= clean_df_severe_flu %>% 
  ggplot(aes(x=month, y= case, group=year))+
  geom_line(color="grey10", size=1, alpha=0.6, linetype="dashed")+
  geom_line(data=clean_df_severe_flu %>% filter(year==2020),
            color="darkred", size=1.2)+
  annotate("rect", xmin = 1, xmax = 6, ymin = 0, ymax = Inf,
           alpha =0.2, fill="pink")+
  annotate("rect", xmin = 10, xmax = 12, ymin = 0, ymax = Inf,
           alpha = 0.2 ,fill="pink")+
  annotate("text", x = 10.7, y = 1050, label = "Flu season: Oct ~ May", size=4)+
  annotate("text", x = 0.7, y = 380, label = "2020", color="darkred", fontface="bold")+
  annotate("text", x = 0.7, y = 250, label = "2019", color="grey10", fontface="bold",alpha=0.6)+
  annotate("text", x = 1.7, y = 1050, label = "2016", color="grey10", fontface="bold",alpha=0.6)+  
  scale_color_aaas()+
  labs(x=NULL, y= "Reported cases",
       title="Yearly comparison between 2010 ~ 2020 monthly number of national cases",
       subtitle = "Disease: severe complication influenza, indigenous and imported",
       caption = "Data source: Taiwan CDC",
       size=10)+
  theme_bw()+
  theme(text=element_text(size=20),
        plot.title = element_text(size=15),
        plot.subtitle = element_text(size=10)
  )


plot = ggplotly(a)

plot[['x']][['layout']][['shapes']] <- c()

plot %>% 
  layout(font=list(family="helvetica"),
         shapes = list(
           list(tyep="rect",
                xref="x", yref="paper",
                x0=1, y0=0, x1=6, y1=1, 
                fillcolor = "pink", line = list(color = "pink"),
                opacity=0.2),
           list(tyep="rect",
                xref="x", yref="paper",
                x0=10, y0=0, x1=12, y1=1, 
                fillcolor = "pink", line = list(color = "pink"),
                opacity=0.2)),
         annotations = 
           list(x = 1, y = -0.1, text = "Data source: Taiwan CDC", 
                showarrow = F, xref='paper', yref='paper', 
                xanchor='right', yanchor='auto', xshift=-30, yshift=-30,
                font=list(size=13, color="grey 10")),
         margin = list(t=100),
         title = list(
           font = list(
             size = 21
           ), 
           text = "Yearly comparison between 2010 ~ 2020 monthly number of national cases <br><sub>Disease: Influenza with severe complications, indigenous and imported</sub>"
         )
         )
```


### Outpatient of influenza
<!-- ## 流感及其所致肺炎門診 -->

```{r, echo=F, fig.align='center', fig.width=9}

df_flu_lung <- read.csv("data/taiwan_cdc/健保流感及其所致肺炎門診年齡別就診人次趨勢圖(2010年01週_2020年18週).csv",stringsAsFactors = )


week_map = read_excel("data/taiwan_cdc/WEEKDATE.xls")

week_date_map = week_map %>% 
  mutate_if(is.character, as.numeric) %>% 
  group_by(year, week) %>% 
  # mutate(week_day = rank(date)) %>% 
  # mutate(max_week_day = max(week_day)) %>% 
  filter(rank(date)== max(rank(date))) %>% 
  ungroup()


clean_column <- function(df){
  list_new_col = df %>% colnames()
  
  list_new_col[1]="week"
  
  list_new_col = str_replace_all(list_new_col, "X","")
  
  colnames(df) = list_new_col
  
  df %>% 
    gather(key="year", value="case", -week) %>%
    mutate_if(is.character, as.numeric) %>% 
    inner_join(week_date_map, by=c("year","week")) %>% 
    mutate(onset_month=month(date)) %>% 
    group_by(year, onset_month) %>% 
    summarise(case = sum(case, na.rm = T)) %>% 
    ungroup() %>% 
    mutate(flu_season = if_else(onset_month>=6 & onset_month<=9,
                                "not flu season",
                                "flu season")) %>%
    mutate(month= factor(month.abb[onset_month], levels = month.abb, ordered = T)) %>%
    mutate(this_year = if_else(year==2020, "this year", "other years")) %>%
    mutate(year2=year) %>% 
    filter(!(year==2020 & onset_month >=5))
  
}

clean_df_flu_lung <- clean_column(df_flu_lung)

a= clean_df_flu_lung %>% 
  ggplot(aes(x=month, y= case, group=year))+
  geom_line(color="grey10", size=1, alpha=0.6, linetype="dashed")+
  geom_line(data=clean_df_flu_lung %>% filter(year==2020),
            color="darkred", size=1.2)+
  annotate("rect", xmin = 1, xmax = 6, ymin = 0, ymax = Inf,
           alpha =0.2, fill="pink")+
  annotate("rect", xmin = 10, xmax = 12, ymin = 0, ymax = Inf,
           alpha = 0.2 ,fill="pink")+
  annotate("text", x = 10.8, y = 420000, label = "Flu season: Oct ~ May", size=4)+
  annotate("text", x = 0.7, y = 285000, label = "2020", color="darkred", fontface="bold")+  
  scale_y_continuous(labels = label_number_si(accuracy = 1.0),
                     breaks = pretty_breaks(5))+
  labs(x=NULL, y= "Reported cases",
       title="Yearly comparison between 2010 ~ 2020 monthly number of national cases",
       subtitle = "Influenza, outpatient",
       caption = "Data source: Taiwan CDC",
       size=10)+
  theme_bw()+
  theme(text=element_text(size=20),
        plot.title = element_text(size=15),
        plot.subtitle = element_text(size=10)
  )

plot = ggplotly(a)

plot[['x']][['layout']][['shapes']] <- c()

plot %>% 
  layout(font=list(family="helvetica"),
         shapes = list(
           list(tyep="rect",
                xref="x", yref="paper",
                x0=1, y0=0, x1=6, y1=1, 
                fillcolor = "pink", line = list(color = "pink"),
                opacity=0.2),
           list(tyep="rect",
                xref="x", yref="paper",
                x0=10, y0=0, x1=12, y1=1, 
                fillcolor = "pink", line = list(color = "pink"),
                opacity=0.2)),
         annotations = 
           list(x = 1, y = -0.1, text = "Data source: Taiwan CDC", 
                showarrow = F, xref='paper', yref='paper', 
                xanchor='right', yanchor='auto', xshift=-15, yshift=-30,
                font=list(size=13, color="grey 10")),
         margin = list(t=100),
         title = list(
           font = list(
             size = 20
           ), 
           text = "Yearly comparison between 2010 ~ 2020 monthly number of national cases <br><sub>Disease: Influenza, outpatient</sub>"
         ))


```


### Outpatient of other Pneumonia

<!-- ## 其他肺炎 -->

```{r, echo=F, fig.align='center', fig.width=9}

df_other_lung = read.csv("data/taiwan_cdc/健保其他肺炎門診年齡別就診人次趨勢圖(2010年01週_2020年18週).csv", stringsAsFactors = F)

clean_df_other_lung = clean_column(df_other_lung)

a= clean_df_other_lung  %>% 
  ggplot(aes(x=month, y= case, group=year))+
  geom_line(color="grey10", size=1, alpha=0.6, linetype="dashed")+
  geom_line(data=clean_df_other_lung  %>% filter(year==2020),
            color="darkred", size=1.2)+
  annotate("rect", xmin = 1, xmax = 6, ymin = 0, ymax = Inf,
           alpha =0.2, fill="pink")+
  annotate("rect", xmin = 10, xmax = 12, ymin = 0, ymax = Inf,
           alpha = 0.2 ,fill="pink")+
  annotate("text", x = 10.8, y = 185000, label = "Flu season: Oct ~ May", size=4)+
  annotate("text", x = 0.7, y = 114000, label = "2020", color="darkred", fontface="bold")+
  scale_y_continuous(labels = label_number_si(accuracy = 1.0),
                     breaks = pretty_breaks(5))+
  labs(x=NULL, y= "Reported cases",
       title="Yearly comparison between 2010 ~ 2020 monthly number of national cases",
       subtitle = "Other Pneumonia",
       caption = "Data source: Taiwan CDC",
       size=10)+
  theme_bw()+
  theme(text=element_text(size=20),
        plot.title = element_text(size=15),
        plot.subtitle = element_text(size=10)
  )


plot = ggplotly(a)

plot[['x']][['layout']][['shapes']] <- c()

plot %>% 
  layout(font=list(family="helvetica"),
         shapes = list(
           list(tyep="rect",
                xref="x", yref="paper",
                x0=1, y0=0, x1=6, y1=1, 
                fillcolor = "pink", line = list(color = "pink"),
                opacity=0.2),
           list(tyep="rect",
                xref="x", yref="paper",
                x0=10, y0=0, x1=12, y1=1, 
                fillcolor = "pink", line = list(color = "pink"),
                opacity=0.2)),
         annotations = 
           list(x = 1, y = -0.1, text = "Data source: Taiwan CDC", 
                showarrow = F, xref='paper', yref='paper', 
                xanchor='right', yanchor='auto', xshift=-15, yshift=-30,
                font=list(size=13, color="grey 10")),
         margin = list(t=100),
         title = list(
           font = list(
             size = 20
           ), 
           text = "Yearly comparison between 2010 ~ 2020 monthly number of national cases <br><sub>Disease: other Pneumonia</sub>"
         ))
```

## Enterovirus

### Enterovirus infection with severe complications
<!-- ## 全國腸病毒感染併發重症本土病例及境外移入 -->
```{r, echo=F, fig.align='center', fig.width=9}
df_ent = read.csv("data/taiwan_cdc/全國腸病毒感染併發重症本土病例及境外移入病例同期比較趨勢圖_all2.csv", stringsAsFactors = F)

clean_column <- function(df){
  list_new_col = df %>% colnames()
  
  list_new_col[1]="onset_month"
  
  list_new_col = str_replace_all(list_new_col, "X","")
  
  colnames(df) = list_new_col
  
  df %>% 
    gather(key="year", value="case", -onset_month) %>% 
    mutate(flu_season = if_else(onset_month>=6 & onset_month<=9,
                                "not flu season",
                                "flu season")) %>% 
    mutate(month= factor(month.abb[onset_month], levels = month.abb, ordered = T)) %>%
    mutate(plot_date = ymd(str_c(year,"-",month,"-","15"))) %>% 
    mutate(this_year = if_else(year==2020, "this year", "other years")) %>% 
    mutate(year2=year) %>% 
    filter(!(year==2020 & onset_month >=5))
}

clean_df_ent = clean_column(df_ent)

a= clean_df_ent %>% 
  ggplot(aes(x=month, y= case, group=year))+
  geom_line(color="grey10", size=1, alpha=0.6, linetype="dashed")+
  geom_line(data=clean_df_ent %>% filter(year==2020),
            color="darkred", size=1.2)+
  annotate("rect", xmin = 1, xmax = 6, ymin = 0, ymax = Inf,
           alpha =0.2, fill="pink")+
  annotate("rect", xmin = 10, xmax = 12, ymin = 0, ymax = Inf,
           alpha = 0.2 ,fill="pink")+
      annotate("text", x = 9.3, y = 37, label = "Enterovirus season: Apr ~ Nov", size=4)+
  annotate("text", x = 0.7, y = 4, label = "2020", color="darkred", fontface="bold")+
  # annotate("text", x = 0.7, y = 250, label = "2019", color="grey10", fontface="bold",alpha=0.6)+
  # annotate("text", x = 1.7, y = 1050, label = "2016", color="grey10", fontface="bold",alpha=0.6)+  
  scale_color_aaas()+
  labs(x=NULL, y= "Reported cases",
       title="Yearly comparison between 2010 ~ 2020 monthly number of national cases",
       caption = "Data source: Taiwan CDC",
       size=10)+
  theme_bw()+
  theme(text=element_text(size=20),
        plot.title = element_text(size=15),
        plot.subtitle = element_text(size=10)
  )

plot = ggplotly(a)

plot[['x']][['layout']][['shapes']] <- c()

plot %>% 
  layout(font=list(family="helvetica"),
         shapes = list(
           list(tyep="rect",
                xref="x", yref="paper",
                x0=4, y0=0, x1=11, y1=1, 
                fillcolor = "pink", line = list(color = "pink"),
                opacity=0.2)),
         annotations = 
           list(x = 1, y = -0.1, text = "Data source: Taiwan CDC", 
                showarrow = F, xref='paper', yref='paper', 
                xanchor='right', yanchor='auto', xshift=-15, yshift=-30,
                font=list(size=13, color="grey 10")),
         margin = list(t=100),
         title = list(
           font = list(
             size = 20
           ), 
           text = "Yearly comparison between 2010 ~ 2020 monthly number of national cases <br><sub>Disease: Enterovirus infection with severe complications, indigenous and imported</sub>"
         ))
```


### Outpatient and emergency of Enterovirus infection
<!-- ## 健保腸病毒門診+急診 -->

```{r, echo=F, fig.align='center', fig.width=9}

df_ent_case = read.csv("data/taiwan_cdc/健保腸病毒門診-+-急診年齡別就診人次趨勢圖(2010年01週_2020年18週).csv", stringsAsFactors = F)

clean_column <- function(df){
  list_new_col = df %>% colnames()
  
  list_new_col[1]="week"
  
  list_new_col = str_replace_all(list_new_col, "X","")
  
  colnames(df) = list_new_col
  
 df %>% 
   gather(key="year", value="case", -week) %>%
   mutate_if(is.character, as.numeric) %>% 
   inner_join(week_date_map, by=c("year","week")) %>% 
   mutate(onset_month=month(date)) %>% 
   group_by(year, onset_month) %>% 
   summarise(case = sum(case, na.rm = T)) %>% 
   ungroup() %>% 
   mutate(flu_season = if_else(onset_month>=6 & onset_month<=9,
                               "not flu season",
                               "flu season")) %>%
   mutate(month= factor(month.abb[onset_month], levels = month.abb, ordered = T)) %>%
   mutate(this_year = if_else(year==2020, "this year", "other years")) %>%
   mutate(year2=year) %>% 
   filter(!(year==2020 & onset_month >=5))

}


clean_df_ent_case = clean_column(df_ent_case) 


a= clean_df_ent_case %>% 
  ggplot(aes(x=month, y= case, group=year))+
  geom_line(color="grey10", size=1, alpha=0.6, linetype="dashed")+
  geom_line(data=clean_df_ent_case %>% filter(year==2020),
            color="darkred", size=1.2)+
  annotate("rect", xmin = 3, xmax = 11, ymin = 0, ymax = Inf,
           alpha =0.2, fill="pink")+
  annotate("text", x = 0.7, y = 22000, label = "2020", color="darkred", fontface="bold")+
  annotate("text", x = 9.3, y = 165000, label = "Enterovirus season: Apr ~ Nov", size=4)+
  scale_color_aaas()+
  scale_y_continuous(labels = label_number_si(accuracy = 1.0),
                     breaks = pretty_breaks(5))+
  labs(x=NULL, y= "Reported cases",
       title="Yearly comparison between 2010 ~ 2020 monthly number of national cases",
       caption = "Data source: Taiwan CDC",
       size=10)+
  theme_bw()+
  theme(text=element_text(size=20),
        plot.title = element_text(size=15),
        plot.subtitle = element_text(size=10)
  )

plot = ggplotly(a)

plot[['x']][['layout']][['shapes']] <- c()

plot %>% 
  layout(font=list(family="helvetica"),
         shapes = list(
           list(tyep="rect",
                xref="x", yref="paper",
                x0=4, y0=0, x1=11, y1=1, 
                fillcolor = "pink", line = list(color = "pink"),
                opacity=0.2)),
         annotations = 
           list(x = 1, y = -0.1, text = "Data source: Taiwan CDC", 
                showarrow = F, xref='paper', yref='paper', 
                xanchor='right', yanchor='auto', xshift=-15, yshift=-30,
                font=list(size=13, color="grey 10")),
         margin = list(t=100),
         title = list(
           font = list(
             size = 20
           ), 
           text = "Yearly comparison between 2010 ~ 2020 monthly number of national cases <br><sub>Disease: Enterovirus infection, outpatient and emergency </sub>"
         ))



```

# Reference 

1. [Taiwan CDC](https://www.cdc.gov.tw/En)
    - [COVID-19](https://sites.google.com/cdc.gov.tw/2019-ncov/taiwan)
    - [Influenza](https://nidss.cdc.gov.tw/en/CDCWNH07.aspx?dc=1&dt=4&disease=487a)
    - [Enterovirus](https://nidss.cdc.gov.tw/en/CDCWNH07.aspx?dc=1&dt=3&disease=0749)
    - [Other Pneumonia](https://nidss.cdc.gov.tw/en/CDCWNH09.aspx)
2. [Post in PTT by TheoEpstein](https://www.ptt.cc/bbs/Gossiping/M.1587879423.A.012.html?fbclid=IwAR2xO8xVmjyuyJxnmLG07489Wm-J7o9qet5vyS4O5d47UuF9E-si5YvLbEI) 
3. [COVID19 cases in Taiwan](https://docs.google.com/spreadsheets/d/1I9EXxe-pWLhcLosakg5TPt98ERY6tdpJn1KngIGY7oY/edit#gid=0)

* All data are downloaded as of May 9, 2020. The data of May 2020 is not included in the plots since the data of May is not yet complete.
